Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers

### 2023-01-04 Lab 14  RDD/Lab14_rdd_Answers.Rmd
Add files via upload

---
title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV)"
author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
output: html_document
bibliography: rdd.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE)
```

@@

# Plot  previous vote margin (x) versus vote for current election (y) by Senate class (color)
ggplot(_____, aes(x = _____, y = _____, color = ______) +
  geom_point() +
  theme_bw() +
  scale_color_discrete(name = 'Senate \n Class', labels = c('1', '2', '3')) +
  geom_vline(xintercept = ____) +
  theme(legend.position = 'right',
        plot.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank())
```

@@

```{r}
# Use Rdplot to see how outcome varies on either side of cutoff
rdplot(senate$____, senate$______)
```

@@

```{r}
# Fit model 
fit_rd <- rdrobust(senate$_____, senate$______)
summary(fit_rd)
```

@@

```{r}
# Use rdrobust to estimtate the model with covs argument specified
fit_rd_covs <- rdrobust(senate$____, senate$______, 
                        covs=_____________________)

@@

```{r}
# Density test
DCdensity(senate$_____, ___)
```

@@

```{r}
#Reading in the data
women<- read.csv("http://eclr.humanities.manchester.ac.uk/images/5/5f/Mroz.csv")

@@

```

# Exercise: Do Voters change the mind of their representatives? Evidence from the U.S. House.

@@

```{r}
# Load Data
enricoall4 <- read_dta("enricoall4.dta")

# Fit RDD Model specifying clustered standard errors for ADA as DV

# Plot RDD

```

```{r}
# Fit RDD Model specifying clustered standard errors for DW-Nom as DV

# Plot RDD

```

@@

```{r}
#McCrary Test

```
The McCrary Test does not show sharp jumps at around the threshold (c=0.5), but it does alert us that there are heavy tails in the distribution of the running variable. The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win). You can also see this above with the lack of observations around the cutpoint. This does lead to some concerns about the robustness of the RD effect we found, and particularly in the sudden reversal of RD trends at the upper tail of the running variable. Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!

# Bibliography

>>>>

---
title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers"
author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```

@@

# Plot  previous vote margin (x) versus vote for current election (y) by Senate class (color)
ggplot(senate, aes(x = margin, y = vote, color = as.factor(class))) +
  geom_point() +
  theme_bw() +
  scale_color_discrete(name = 'Senate \n Class', labels = c('1', '2', '3')) +
  geom_vline(xintercept = 0) +
  theme(legend.position = 'right',
        plot.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank())
```

@@

```{r}
# Use Rdplot to see how outcome varies on either side of cutoff
rdplot(senate$vote, senate$margin)
```

@@

```{r}
# Fit model 
fit_rd <- rdrobust(senate$vote, senate$margin)
summary(fit_rd)
```

@@

```{r}
# Use rdrobust to estimtate the model with covs argument specified
fit_rd_covs <- rdrobust(senate$vote, senate$margin, covs=senate[,c("termssenate","termshouse","class")])

@@

```{r}
# Density test
DCdensity(senate$margin, 0)
```

@@

```{r}

#Reading in the data
women<- read.csv("http://eclr.humanities.manchester.ac.uk/images/5/5f/Mroz.csv")

@@

```
# Exercise: Do Voters change the mind of their representatives? Evidence from the U.S. House.

@@

```{r}
# Load Data
enricoall4 <- read_dta("enricoall4_small.dta")

# Fit RDD Model specifying clustered standard errors for ADA as DV
lmb.fit <- rdrobust(enricoall4$realada,
                    enricoall4$lagdemvoteshare,
                    cluster=enricoall4$clusterid,
                    c=0.5)

# Plot RDD
rdplot(enricoall4$realada,
       enricoall4$lagdemvoteshare,
       x.label="Democrat Vote Share at time t",
       y.label="ADA Score at time t+1",
       c=0.5)
```

```{r}
# Fit RDD Model specifying clustered standard errors for DW-Nom as DV
lmb.fit2 <- rdrobust(enricoall4$dwnom1,
                     enricoall4$lagdemvoteshare,
                     cluster=enricoall4$clusterid,
                     c=0.5)
# Plot RDD
rdplot(enricoall4$dwnom1,
         enricoall4$lagdemvoteshare,
       x.label="Democrat Vote Share at time t",
       y.label="DW-Nominate Score at time t",
         c=0.5)
```

@@

```{r}
#McCrary Test
DCdensity(enricoall4$lagdemvoteshare, 0.5)
```
The McCrary Test does not show sharp jumps at around the threshold (c=0.5), but it does alert us that there are heavy tails in the distribution of the running variable. The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win). You can also see this above with the lack of observations around the cutpoint. This does lead to some concerns about the robustness of the RD effect we found, and particularly in the sudden reversal of RD trends at the upper tail of the running variable. Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!
